Faster generic IND-CCA2 secure KEM using ``encrypt-then-MAC''

### 2024-10-04 etm-transform/manuscript/main.tex
Do not commit references in pdf

It is difficult to build a provably IND-CCA2 secure KEM from scratch. Instead, secure KEMs are usually built on top of a PKE with weaker security property (e.g. being only OW-CPA or IND-CPA secure, not IND-CCA2). One such construction is the Fujisaki-Okamoto transformation, proposed in 1999 by Fujisaki Eiichiro and Okamoto Tetsuyaki in their seminal paper \cite{DBLP:conf/crypto/FujisakiO99}\cite{DBLP:journals/joc/FujisakiO13}. The first Fujisaki-Okamoto transformation combines an OW-CPA secure PKE with an IND-CPA secure symmetric cipher into a hybrid PKE (HPKE) with proven IND-CCA2 security under the random oracle model. Subsequent works \cite{DBLP:conf/ctrsa/OkamotoP01}\cite{DBLP:conf/ctrsa/CoronHJPPT02}\cite{DBLP:conf/ima/Dent03} improved on the original proposal and adapted it to build KEM instead of HPKE. This line of work culminated in a landmark publication in 2017 by Hofhein, Hovelmann, and Kiltz \cite{DBLP:conf/tcc/HofheinzHK17}\cite{DBLP:conf/asiacrypt/HovelmannsHM22}, where the authors provided a versatile variety of modular KEM constructions with tight security reduction in the random oracle model and non-tight security reduction in the quantum random oracle model.

@@

Where the expectation is taken with respect to the probability distribution of all possible keypairs $(\pk, \sk) \leftsample \texttt{PKE.KeyGen()}$. For many lattice-based cryptosystems, including ML-KEM, decryption failures could leak information about the secret key, although the probability of a decryption failure is low enough that classical adversaries cannot exploit decryption failure more than they can defeat the underlying lattice problem. On the other hand, a quantum adversary may be able to exploit decryption failure in reasonable runtime by efficiently searching through all possible inputs using Grover's search algorithm. For that, ML-KEM made slight modifications in its KEM construction to prevent quantum adversary from precomputing large lookup table. We refer readers to \cite{avanzi2019crystals} and \cite{bos2018crystals} for the details.

@@

\subsection{Key exchange protocols}\label{sec:key-exchange-protocols}
A common application of key encapsulation mechanism is key exchange protocols, where two parties establish a shared secret using a public channel. \cite{bos2018crystals} described three key exchange protocols: unauthenticated key exchange (KE), unilaterally authenticated key exchange (UAKE), and mutually authenticated key exchange (AKE). We instantiated an implementation for each of the three key exchange protocols using different variations of the ``encrypt-then-MAC'' KEM and compared round trip time with implementations instantiated using ML-KEM.

>>>>

It is difficult to build a provably IND-CCA2 secure KEM from scratch. Instead, secure KEMs are usually built on top of a PKE with weaker security property (e.g. being only OW-CPA or IND-CPA secure, not IND-CCA2). One such construction is the Fujisaki-Okamoto transformation, proposed in 1999 by Fujisaki Eiichiro and Okamoto Tetsuyaki in their seminal paper \cite{DBLP:conf/crypto/FujisakiO99}\cite{DBLP:journals/joc/FujisakiO13}. The first Fujisaki-Okamoto transformation combines an OW-CPA secure PKE with an IND-CPA secure symmetric cipher into a hybrid PKE (HPKE) with proven IND-CCA2 security under the random oracle model. Subsequent works \cite{DBLP:conf/ctrsa/OkamotoP01}\cite{DBLP:conf/ctrsa/CoronHJPPT02}\cite{DBLP:conf/ima/Dent03} improved on the original proposal and adapted it to build KEM instead of HPKE. This line of work culminated in a landmark publication in 2017 by Hofhein, Hovelmann, and Kiltz \cite{hofheinz2017modular}\cite{DBLP:conf/asiacrypt/HovelmannsHM22}, where the authors provided a versatile variety of modular KEM constructions with tight security reduction in the random oracle model and non-tight security reduction in the quantum random oracle model.

@@

Where the expectation is taken with respect to the probability distribution of all possible keypairs $(\pk, \sk) \leftsample \texttt{PKE.KeyGen()}$. For many lattice-based cryptosystems, including ML-KEM, decryption failures could leak information about the secret key, although the probability of a decryption failure is low enough that classical adversaries cannot exploit decryption failure more than they can defeat the underlying lattice problem. On the other hand, a quantum adversary may be able to exploit decryption failure in reasonable runtime by efficiently searching through all possible inputs using Grover's search algorithm. For that, ML-KEM made slight modifications in its KEM construction to prevent quantum adversary from precomputing large lookup table. We refer readers to \cite{avanzi2019crystals} and \cite{DBLP:conf/eurosp/BosDKLLSSSS18} for the details.

@@

\subsection{Key exchange protocols}\label{sec:key-exchange-protocols}
A common application of key encapsulation mechanism is key exchange protocols, where two parties establish a shared secret using a public channel. \cite{DBLP:conf/eurosp/BosDKLLSSSS18} described three key exchange protocols: unauthenticated key exchange (KE), unilaterally authenticated key exchange (UAKE), and mutually authenticated key exchange (AKE). We instantiated an implementation for each of the three key exchange protocols using different variations of the ``encrypt-then-MAC'' KEM and compared round trip time with implementations instantiated using ML-KEM.
